---
title: "P8105 HW 1"
author: Jennifer Mizhquiri
output: github_document
---


The following package will be necessary:

```{r library, message = FALSE}
library(tidyverse)

```

# Problem 1 

```{r load_penguins, echo = FALSE}

data("penguins", package = "palmerpenguins")

head(penguins)

```


```{r, echo = FALSE}
flipper_length_pull <- penguins %>% pull(flipper_length_mm) 


bill_length_pull <- penguins %>% pull(bill_length_mm) 

```

A preview of the _penguins_ dataset is displayed above. It contains geographic and demographic data on individual penguins belonging to different species of penguins. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns: species (species), island (island), bill_length_mm (bill length in mm), bill_depth_mm (bill depth in mm), flipper_length_mm (flipper length in mm), body_mass_g (body mass in grams), sex (sex), and year (year). From here, summary statistics could be extracted. The mean flipper length is `r mean((flipper_length_pull), na.rm = TRUE)` mm. The mean bill length is `r mean((bill_length_pull), na.rm = TRUE)` mm.


 _Scatterplot Comparing Bill and Flipper Lengths in Penguins_
```{r penguins_scatterplot, echo = FALSE, message = FALSE, warning = FALSE}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper length (mm)") + ggtitle("Comparing Penguin Flipper and Bill Lengths")

ggsave("BillxFlipperLengthxSpecies.PDF", height = 4, width = 10)

```

On average, Gentoo sp. have longer bills and flipper lengths than Adelie and longer flippers and approximately equal bill lengths as the Chinstrap species. Adelie species have shorter bill lengths and smaller flipper lengths on average relative to the Chinstrap species (though outliers are present and the average difference does not appear large). Chinstrap have bill lengths about as  long as the Gentoo species but shorter flipper lengths. 

# Problem 2

```{r sample_df, message = FALSE, echo = FALSE}
sample_df = 
  tibble(
  var_rand = rnorm(n = 10,mean = 0),
  var_logic = var_rand > 0, 
  var_char = c("a", "b","c","a", "a", "b", "b", "c", "b","a"),
  var_factor = factor(var_char, levels = c("a","b","c"), ordered = TRUE)
  )
```


```{r, echo = FALSE}
print(sample_df)
```

Above is a dataframe comprised of: 

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”
  * Note: this factor vector is a factor version of the character version, but has assigned levels to each of the possible values

 _Below is a code chunk demonstrating how variables may be pulled._
```{r pulling_var}

var_rand_pull <- sample_df %>% pull(var_rand) 
var_logic_pull <- sample_df %>% pull(var_logic)
var_char_pull <- sample_df %>% pull(var_char)
var_factor_pull <- sample_df %>%  pull(var_factor)
```
 
 _What happens when you take the mean of each variable?_
 
* The mean of var_rand is `r mean(var_rand_pull)` 
  * It makes sense that there is a mean as the variable is a numeric one. The mean is expected to also be close to 0. 
* The mean of var_logic is `r mean(var_logic_pull)` 
  * It makes sense that there is an available mean as R interprets "True" as 1 and "False" as 0, indicating that with a number closer to 0 there are more instances of "False" and vice versa.
* The mean of var_char is `r mean(var_char_pull)`. 
  * It makes sense that there is no available mean as the variable is that of a character and as such it is not applicable. 
* The mean of var_factor is `r mean(var_factor_pull)`
  * It makes sense that there is no available mean as the variables are made up of levels but even so it is not numeric. 
 
_What happens when you try to explicitly convert a factor, character, and logical character to a numeric one?_

 
```{r coercion, message = FALSE, results = FALSE, warning = FALSE}

var_char_num <- as.numeric(var_char_pull)
var_factor_num <- as.numeric(var_factor_pull)
var_logic_num <- as.numeric(var_logic_pull)


```
 

R will provide a warning message noting that NAs are introduced by coercion for a character variable. 
  
  * This makes sense as R does not have a basis for ascribing a numeric value to the characters, which is why you still yield an NA, and furthermore why it was not possible (nor logical in this instance) to have a mean. Rather, the mean appeared as NA.

No such warning messages appeared for the formerly logical and factor variables. 
  
  * In the case of the logical variable, I would expect the same mean to appear (a numeric mean was produced before the logical variable was coerced into numeric, as described above. In the case of the factor variable, the coercion to numeric variable meant that each of the levels of 1,2,3 were interpreted as numeric and the average of that collection would now be able to be calculated.
  

