---
title: "P8105 HW 1"
author: Jennifer Mizhquiri
output: github_document
---


# Problem 1 

Load packages (below) & installed _penguins_ dataset (not shown)

```{r library, message = FALSE}
library(tidyverse)
library(dplyr)
```


```{r load_penguins}

data("penguins", package = "palmerpenguins")

head(penguins)

```

A preview of the _penguins_ dataset is displayed above. It contains geographic and demographic data on individuals belonging to different species of penguins. There are `r nrow(penguins)` records and `r ncol(penguins)` variables: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year. 

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm




```{r penguins_scatterplot, message = FALSE, warning = FALSE}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper length (mm)") + ggtitle("Gentoo sp. have longer bills and flipper lengths than Adelie and Chinstrap species")

ggsave("BillxFlipperLengthxSpecies.PDF", height = 4, width = 10)

```


# Problem 2

```{r sample_df, message = FALSE}
sample_df = 
  tibble(
  var_rand = rnorm(n = 10,mean = 0),
  var_logic = var_rand > 0, 
  var_char = c("a", "b","c","d", "a", "b", "b", "c", "d","a"),
  var_factor = factor(var_char, levels = c("a","b","c", "d"), ordered = TRUE)
  )
```


```{r}
print(sample_df)
```
 Above is a dataframe comprised of: 
- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r pulling_var}

var_rand_pull <- sample_df %>% pull(var_rand) #is it best practice to name a pulled variable or simply reuse? It's okay to reuse but could rename something else
var_logic_pull <- sample_df %>% pull(var_logic) #why is this providing an answer?
var_char_pull <- sample_df %>% pull(var_char)
var_factor_pull <- sample_df %>%  pull(var_factor)
```
 The mean of var_rand is `r mean(var_rand_pull)`
 The mean of var_logic is `r mean(var_logic_pull)`
 The mean of var_char is `r mean(var_char_pull)`
 The mean of var_factor is `r mean(var_factor_pull)`
 
 
```{r coercion}

var_char_num <- as.numeric(var_char_pull)
var_factor_num <- as.numeric(var_factor_pull)
mean(var_factor_num)
mean(var_char_num)

```
 
 

