---
title: "P8105 HW 1"
author: Jennifer Mizhquiri
output: github_document
---


# Problem 1 

Load packages (below) & installed _penguins_ dataset (not shown)

```{r library, message = FALSE}
library(tidyverse)
library(dplyr)
```


```{r load_penguins}

data("penguins", package = "palmerpenguins")

head(penguins)

```

A preview of the _penguins_ dataset is displayed above. It contains geographic and demographic data on individuals belonging to different species of penguins. There are `r nrow(penguins)` records and `r ncol(penguins)` variables: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year. 

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm



```{r penguins_scatterplot, message = FALSE, warning = FALSE}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper length (mm)") + ggtitle("Gentoo sp. have longer bills and flipper lengths than Adelie and Chinstrap species")

ggsave("BillxFlipperLengthxSpecies.PDF", height = 4, width = 6)

```


