---
title: "P8105 HW 1"
author: Jennifer Mizhquiri
output: github_document
---


# Problem 1 

Load packages (below) & installed _penguins_ dataset (not shown)

```{r library, message = FALSE}
library(tidyverse)
library(dplyr)
```


```{r load_penguins}

data("penguins", package = "palmerpenguins")

head(penguins)

```

A preview of the _penguins_ dataset is displayed above. It contains geographic and demographic data on individuals belonging to different species of penguins. There are `r nrow(penguins)` records and `r ncol(penguins)` variables: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year. 

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm




```{r penguins_scatterplot, message = FALSE, warning = FALSE}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper length (mm)") + ggtitle("Comparing Penguing Flipper and Bill Lengths")

ggsave("BillxFlipperLengthxSpecies.PDF", height = 4, width = 10)

```

On average, Gentoo sp. have longer bills and flipper lengths than Adelie and Chinstrap species. Adelie species have shorter bill lengths and smaller flipper lengths on average relative to the Chinstrap species (though outliers are present and the average difference does not appear large). Chinstrap have bill lengths about as  long as the Gentoo species but shorter flipper lengths. 

# Problem 2

```{r sample_df, message = FALSE}
sample_df = 
  tibble(
  var_rand = rnorm(n = 10,mean = 0),
  var_logic = var_rand > 0, 
  var_char = c("a", "b","c","a", "a", "b", "b", "c", "b","a"),
  var_factor = factor(var_char, levels = c("a","b","c"), ordered = TRUE)
  )
```


```{r}
print(sample_df)
```

Above is a dataframe comprised of: 

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r pulling_var}

var_rand_pull <- sample_df %>% pull(var_rand) 
var_logic_pull <- sample_df %>% pull(var_logic)
var_char_pull <- sample_df %>% pull(var_char)
var_factor_pull <- sample_df %>%  pull(var_factor)
```
 
 _What happens when you take the mean of each variable?_
 
* The mean of var_rand is `r mean(var_rand_pull)` 
  * It makes sense that there is a mean as the variable is a numeric one. The mean is expected to also be close to 0. 
* The mean of var_logic is `r mean(var_logic_pull)` 
  * It makes sense that there is an available mean as R interprets "True" as 1 and "False" as 0, indicating that with a mean of 0.3, there are more instances of "False".
* The mean of var_char is `r mean(var_char_pull)`. 
  * It makes sense that there is no available mean as the variable is that of a character and as such it is not applicable. 
* The mean of var_factor is `r mean(var_factor_pull)`
  * It makes sense that there is no available mean as the variables are made up of levels but even so it is not numeric. 
 
 
```{r coercion, message = FALSE}

var_char_num <- as.numeric(var_char_pull)
var_factor_num <- as.numeric(var_factor_pull)


```
 
_What happens when you try to take the means of a character and factor variable, respectively, by setting them to numeric variables?_

 
The mean of a factor variable by coercion to a numeric variable is `r mean(var_factor_num)`. 
  * Note: R has interpreted the levels as 1, 2, 3 and has now taken the average of each instance.


The mean of a character variable by coercion to a numeric variable is `r mean(var_char_num)`
  * Note: R does not have a basis for ascribing a numeric value to the characters, which is why you still yield an NA.
